---
title: "HW1: Power Analysis"
output: html_document
---
```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
Task: estimate the mean number of miles communted in the Eugene-Springfield area. 
Randomly chosen selection of people are asked where they live and how far they have to travel to get from home to work
Figure out how many people should we ask? To acheive reasonalble precision. 
Mean number of miles traveled 
Do people that live in one city travel significantly farther than people living in the other city?

Do a power analysis and include graphs that convey

How accurately the mean number of miles can be estimated

How liekly it is that a real difference in mean between Eugene and Springfield will be detected as a function of N


Power analysus is used to calculate the minimum sample size required so that one can be reasonably likely to detect an effect of a given size

OR

Power analysis can be used to calculate the minimum effect size that is likely to be detected in a study using a given sample size. 

Rule of Thumb

The sample size n of a two-sample ttest with power 80% (Beta=0.2) and significance level alpha = 0.05 should be

n = 16(s^2/d^2)
s is population variance and d=mu_1-mu_2, the to me detected difference in the mean values of both samples 

Do power analysis using simulation

Assume that the distribution of distances traveled in Exponential - use rexp()
```{r}
Edistances <- rexp(10000)+1
Sdistances <- rexp(10000)+3
```
``` {r}
hist(Edistances)
```
```{r}
N <- c(3:10,20,50,100,500,1000)
Etmean <- mean(Edistances)
Stmean <- mean(Sdistances)
Emeanerror <- {}
Smeanerror <- {}
upperE <- {}
lowerE <- {}
upperS <- {}
lowerS <- {}
pv <- {}
t <- {}
for (p in seq_along(N)){
  Esmean<-replicate(100,{
    Egn <- mean(sample(Edistances,N[p]))
  })
  upperE[p] <- mean(Esmean[97:98])
  lowerE[p] <- mean(Esmean[2:3])
  Ssmean<-replicate(100, {
    Spf <- mean(sample(Sdistances,N[p]))
  })
  upperS[p] <- mean(Ssmean[97:98])
  lowerS[p] <- mean(Ssmean[2:3])
  Emeanerror[p] <- mean(abs(Esmean-Etmean))
  Smeanerror[p] <- mean(abs(Ssmean-Stmean))
  
  tt<-t.test(sample(Edistances,N[p]), sample(Sdistances,N[p]))
  t[p] <- tt$statistic
  pv[p] <- tt$p.value
}
pdf("meanError.pdf")
plot(N, Emeanerror, type='l', xlab="sample size (number of people surveyed)", ylab="mean error (miles)", main="Mean error of estimates of distance commuted", ylim=c(0,0.5))
lines(N, Smeanerror, col="blue")
legend("topright", legend=c("Eugene", "Springfield"), col=c("black","blue"), lty=1)
dev.off()
pdf("ttest.pdf")
par(mfrow=c(1,2), mar=c(5,3,4,0.5))
plot(N, pv, type="b", log="x",xlab="", ylab="")
title(ylab="p value", line=2)
abline(h=0.05, col="red", lty=2)
plot(N,t, type="b", log="x",xlab="", ylab="")
title(ylab="t statistic", line=2)
title(main="results of t test between Eugene and Springfield commuted distances", outer=T, line=-2)
title(xlab="sample size (number of people surveyed)", outer=T, line=-2)
dev.off()

pdf("CIs.pdf")
plot(, type='l')
for (p in seq_along(N)){
  lines(c(lowerE[p], upperE[p]),c(p,p))
}
```
The population of Springfield is 62,729
The population of Eugene is 170,457
